---
title: "DAC Demo - Rmarkdown"
author: "Emmanuel Nartey"
date: "2024-02-27"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

# Data Analysis Challenge (DAC) - Spring 2024
In this Rmarkdown file, we will demonstrate how to read in datasets directly from Github and follow that with some explorations.

# Import Libraries
These are some packages you could consider using for your analysis based on what you plan to do.

    tidyverse - collection of packages for diverse tasks which include data manipulation (dplyr), reading and writing data (readr), and creating static data visualizations (ggplots).
    Plotly - for intereactive data visualization
    stats - for statistical analysis
    


```{r}
#Install the packages first before import.

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(stats))
suppressPackageStartupMessages(library(knitr))
```



# Data Exploration
```{r}
df <- read_csv("https://raw.githubusercontent.com/cmich-stats-club/spring2024-dac/main/datasets/cbb.csv", show_col_types = FALSE)
```


```{r}
str(df)
```

There are 2,455 records across 24 columns in the dataset.   

```{r}
summary(df)
```

From the YEAR summaries, we see that this data is the seasons 2013 through 2019.


```{r}
#Get the number of missing values for each variable
sapply(df, function(x) sum(is.na(x)))
```

The only columns with missing records are POSTSEASON and SEED. Are these missing on purpose or it is a data anomaly?  

```{r}
#Print out first 6 records
df %>%
  head(.) %>%
  kable(.)
```



```{r}
length(unique(df$TEAM)); length(unique(df$CONF))
```

There are 355 teams (schools) spanning 35 conferences.  

## NCAA Post-Season
```{r}
#I want to focus on only the teams that made it to the post season
df_post_season <- df %>%
  filter(!is.na(POSTSEASON))
```


```{r}
#Number of records in the post season subset table
dim(df_post_season)
```



```{r}
#Which conferences had the most representation post season 2013-2019
df_post_season %>%
  count(CONF) %>%
  arrange(desc(n))  %>%
  kable(.)
```


The ACC is the most represented conference, followed by the Big 10 (B10).  
CMU is in the MAC. This conference has one representative per season. Who are these MAC representatives in the March Madness?  


```{r}
df_post_season %>%
  filter(CONF == "MAC") %>%
  arrange(desc(YEAR)) %>%
  kable(.)
```


Buffalo is the most dominant team in the MAC, 4 March Madness appearances in 7 years. CMU need to improve their odds of making it to the NCAA tournament.  
Unfortunately, the farthest the MAC teams have gone is the Round of 32.

